---
title: "Figures"
author: "Ulrich Lösener"
date: "8 5 2022"
output: html_document
---

This file contains all the code necessary to produce the figures in the paper. 
Note that first, the models need to be run in the markdown file "Formal_models1&2".

### Beta distribution 
The distribution from with health parameters are drawn for each individual can be
visualised with the following code.

```{r Beta Distribution}
# define range (from 0 to 1)
range <- seq(0, 1, length=100)

# create plot of Beta distribution with shape parameters 1.3 and 6
plot(range, dbeta(range, 1.3, 6), type='l', lwd=2, bty="n", xlab="x", ylab="Density", main="Beta Distribution (Shape = 1.3, 6)")
```


### Data generation model 1
Without intervention - 15 days
```{r}
par(mfrow=c(2,3))
plot(ppl.long[511:526, 2], type="b", ylab="Trigger", xlab="Days", main="Trigger", bty="n")
plot(ppl.long[511:526, 3], type="b", ylab="Anxiety", main="Anxiety", xlab="Days", bty="n")
plot(ppl.long[511:526, 4], type="b", ylab="RSB", main="Restorative SB", xlab="Days", bty="n")
plot(ppl.long[511:526, 5], type="b", ylab="Belief", main="Negative Belief", xlab="Days", bty="n")
plot(ppl.long[511:526, 6], type="b", ylab="PSB", main="Preventive SB", xlab="Days", bty="n")
```

With intervention on day 15 - 30 days
```{r}
par(mfrow=c(2,3))
plot(ppl.long[511:541, 2], type="b", ylab="Trigger", xlab="Days", main="Trigger", bty="n")
plot(ppl.long[511:541, 3], type="b", ylab="Anxiety", main="Anxiety", xlab="Days", bty="n")
plot(ppl.long[511:541, 4], type="b", ylab="RSB", main="Restorative SB", xlab="Days", bty="n")
abline(v=15, col="blue")
plot(ppl.long[511:541, 5], type="b", ylab="Belief", main="Negative Belief", xlab="Days", bty="n")
plot(ppl.long[511:541, 6], type="b", ylab="PSB", main="Preventive SB", xlab="Days", bty="n")
abline(v=15, col="blue")
```

### Establish diagnostic criteria
If prevalences over time are to be plotted, comment in lines 50 - 52, 54, 94-98, and comment out line 55.
Otherwise, a single point prevalence is computed
```{r Criteria Model 1}
#sequt <- rev(seq(from=1, to=tpoints, by=30))
#prev <- rep(NA, length(sequt))
#for (i in sequt){
# time frame in which symptoms are measured for diagnosis: 1 if all the time
#t <- i 
t <- 1
# unreasonable, excessive fear 
fearsum <- rep(NA, N)
for(p in 1:N){
  fearsum[p] <- sum(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),3])
}

anx2.crit <- rep(NA, N)
for(p in 1:N){
  anx2.crit[p] <- length(which(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),3]>=2))
}

peakanx <- rep(NA, N)
for(p in 1:N){
  peakanx[p] <- length(which(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),3]==3))
}

# Frequency of triggers
trigsum <- rep(NA, N)
for(p in 1:N){
  trigsum[p] <- sum(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),2])
}

# avoidance or extreme distress
rsbsum <- rep(NA, N)
for(p in 1:N){
  rsbsum[p] <- sum(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),4])
}

psbsum <- rep(NA, N)
for(p in 1:N){
  psbsum[p] <- sum(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),6])
}

# high negative belief towards stimulus
belsum <- rep(NA, N)
for(p in 1:N){
  belsum[p] <- sum(ppl.long[(sequ[p]):(sequ[p]+(tpoints-t)),5])
}
#ppl.wide.prelim <- as.data.frame(cbind(trigsum, fearsum, peakanx, rsbsum, psbsum, belsum))
#ppl.wide.prelim$health <- ifelse(((ppl.wide.prelim$peakanx >= (tpoints-i)/6 & #ppl.wide.prelim$rsbsum >= (tpoints-i)/10) | (ppl.wide.prelim$peakanx >= (tpoints-i)/6 & #ppl.wide.prelim$psbsum >= (tpoints-i)/3)) & ppl.wide.prelim$belsum >= (tpoints-i), 0, 1)
#prev[match(i, sequt)] <- length(which(ppl.wide.prelim$health==0))/N
#}

# which ppl have "healthy" parameter?
par.health <- rep(NA, N)
for(p in 1:N){
par.health[p] <- length(which(ppl.long[(sequ[p]):(sequ[p]+(tpoints-1)),7]==.1))
}
par.health[par.health==tpoints] <- 1

# add diagnostic info in "ppl" matrix
ppl.wide <- as.data.frame(cbind(trigsum, fearsum, anx2.crit, peakanx, rsbsum, psbsum,                                       belsum, par.health))

# assign 1 to healthy ppl and 0 to unhealthy according to the 3 criteria
tpoints.diag <- 360 # time frame for the diagnosis
ppl.wide$health <- ifelse(((ppl.wide$anx2.crit >= tpoints.diag/6 
                            & ppl.wide$rsbsum >= tpoints.diag/10) 
                            | (ppl.wide$anx2.crit >= tpoints.diag/6 
                            & ppl.wide$psbsum >= tpoints.diag/3)) 
                            & ppl.wide$belsum >= tpoints.diag, 0, 1)

# how many meet all criteria?
length(which(ppl.wide$health==0))/N

# combinations of health parameter & actual health
tab <- table(ppl.wide$par.health, ppl.wide$health)
colnames(tab) <- c("unhealthy", "healthy")
rownames(tab) <- c("par unhealthy", "par healthy")
tab
```

Again, if prevalences over time are to be plotted, comment in lines 131 - 133, 135, 176-179, and comment out line 137
Otherwise, a single point prevalence is computed
```{r Criteria Model 2}
#sequt2 <- rev(seq(from=1, to=tpoints2, by=30))
#prev2 <- rep(NA, length(sequt2))
#for (i in sequt2){
# time frame in which symptoms are measured for diagnosis: 1 if all the time
#t <- i

t <- 1

trigsum2 <- rep(NA, N2)
for(p in 1:N2){
 trigsum2[p] <- sum(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),2])
}

# unreasonable, excessive fear 
fearsum2 <- rep(NA, N2)
for(p in 1:N2){
  fearsum2[p] <- sum(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),3])
}

anx2.crit2 <- rep(NA, N2)
for(p in 1:N2){
anx2.crit2[p] <- length(which(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),3]>=2))
}

peakanx2 <- rep(NA, N2)
for(p in 1:N2){
peakanx2[p] <- length(which(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),3]==3))
}

# avoidance or extreme distress 
rsbsum2 <- rep(NA, N2)
for(p in 1:N2){
  rsbsum2[p] <- sum(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),4])
}

psbsum2 <- rep(NA, N2)
for(p in 1:N2){
  psbsum2[p] <- sum(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),6])
}

# high negative belief towards stimulus
belsum2 <- rep(NA, N2)
for(p in 1:N2){
  belsum2[p] <- sum(ppl.long2[(sequ2[p]):(sequ2[p]+(tpoints2-t)),5])
}
#ppl.wide.prelim2 <- as.data.frame(cbind(trigsum2, fearsum2, peakanx2, rsbsum2, psbsum2, belsum2))
#ppl.wide.prelim2$health <- ifelse(((ppl.wide.prelim2$peakanx >= (tpoints2-i)/6 & ppl.wide.prelim2$rsbsum >= (tpoints2-i)/10) | (ppl.wide.prelim2$peakanx >= (tpoints2-i)/6 & ppl.wide.prelim2$psbsum >= (tpoints2-i)/3)) & ppl.wide.prelim2$belsum >= (tpoints2-i), 0, 1)
#prev2[match(i, sequt2)] <- length(which(ppl.wide.prelim2$health==0))/N2
#}

# add diagnostic info in "ppl" matrix
ppl.wide2 <- as.data.frame(cbind(trigsum2, fearsum2, anx2.crit2, peakanx2, rsbsum2, psbsum2, belsum2))

# extract lat parameter for every person
lats2 <- rep(NA, N2)
for (p in 1:N2){
  lats2[p] <- ppl.long2[sequ2[p],7]
}
ppl.wide2$lat2 <- lats2

# add health status for every person according to criteria
tpoints.diag2 <- 360 # set time frame for diagnosis
ppl.wide2$health <- ifelse(((ppl.wide2$anx2.crit2 >= tpoints.diag2/6 & ppl.wide2$rsbsum2 >=        tpoints.diag2/10) | (ppl.wide2$anx2.crit2 >= tpoints.diag2/6 & ppl.wide2$psbsum2 >= tpoints.diag2/3)) & ppl.wide2$belsum2 >= tpoints.diag2, 0, 1)

# how many meet all criteria?
length(which(ppl.wide2$health==0))/N2

# combinations of health parameter & actual health
tab2 <- table((ppl.wide2$lat2<.5), (ppl.wide2$health))
colnames(tab2) <- c("unhealthy", "healthy")
rownames(tab2) <- c("lat2 unhealthy", "lat2 healthy")
tab2
```

### Plot longitudinal data of anxiety and beilef for both models

```{r Time Series Anxiety & Belief Both Models}

# add health info to every person 
health.long <- rep(ppl.wide$health, each=tpoints)
ppl.long <- cbind(ppl.long, health.long)
health.long2 <- rep(ppl.wide2$health, each=tpoints2)
ppl.long2 <- cbind(ppl.long2, health.long2)

# split long data frame of model 1 into healthy and unhealthy
ppl.long.h <- ppl.long[which(ppl.long[,"health.long"]==1),]
ppl.long.p <- ppl.long[which(ppl.long[,"health.long"]==0),]
# same for model 2
ppl.long.h2 <- ppl.long2[which(ppl.long2[,"health.long2"]==1),]
ppl.long.p2 <- ppl.long2[which(ppl.long2[,"health.long2"]==0),]


# placeholders for mean anxiety and belief values model 1
mean.anx <- rep(NA, tpoints)
mean.anx.h <- rep(NA, tpoints)
mean.anx.p <- rep(NA, tpoints)
mean.bel <- rep(NA, tpoints)
mean.bel.h <- rep(NA, tpoints)
mean.bel.p <- rep(NA, tpoints)
# same for model 2
mean.anx2 <- rep(NA, tpoints2)
mean.anx.h2 <- rep(NA, tpoints2)
mean.anx.p2 <- rep(NA, tpoints2)
mean.bel2 <- rep(NA, tpoints2)
mean.bel.h2 <- rep(NA, tpoints2)
mean.bel.p2 <- rep(NA, tpoints2)

# sequences for for loop 
sequ.h2 <- seq(from=1, to=(length(unique(ppl.long.h2[,"ID"]))*tpoints2), by=tpoints2)
sequ.p2 <- seq(from=1, to=(length(unique(ppl.long.p2[,"ID"]))*tpoints2), by=tpoints2)
sequ.h <- seq(from=1, to=(length(unique(ppl.long.h[,"ID"]))*tpoints), by=tpoints)
sequ.p <- seq(from=1, to=(length(unique(ppl.long.p[,"ID"]))*tpoints), by=tpoints)

# store mean anxiety and belief values for healthy and phobics separately - model 1
for (j in 1:tpoints){
  mean.anx[j] <- mean(ppl.long[(sequ-1+j),"anx"])
  mean.anx.h[j] <- mean(ppl.long.h[(sequ.h-1+j),"anx"])
  mean.anx.p[j] <- mean(ppl.long.p[(sequ.p-1+j),"anx"])
  
  mean.bel[j] <- mean(ppl.long[(sequ-1+j),"bel"])
  mean.bel.h[j] <- mean(ppl.long.h[(sequ.h-1+j),"bel"])
  mean.bel.p[j] <- mean(ppl.long.p[(sequ.p-1+j),"bel"])
}
# store mean anxiety and belief values for healthy and phobics separately - model 2
for (j in 1:tpoints2){
  mean.anx2[j] <- mean(ppl.long2[(sequ2-1+j),"anx2"])
  mean.anx.h2[j] <- mean(ppl.long.h2[(sequ.h2-1+j),"anx2"])
  mean.anx.p2[j] <- mean(ppl.long.p2[(sequ.p2-1+j),"anx2"])
  
  mean.bel2[j] <- mean(ppl.long2[(sequ2-1+j),"bel2"])
  mean.bel.h2[j] <- mean(ppl.long.h2[(sequ.h2-1+j),"bel2"])
  mean.bel.p2[j] <- mean(ppl.long.p2[(sequ.p2-1+j),"bel2"])
}

# bind them together - model 2
mean.anx.all <- cbind(mean.anx, mean.anx.h, mean.anx.p)
mean.bel.all <- cbind(mean.bel, mean.bel.h, mean.bel.p)
# and model 2
mean.anx.all2 <- cbind(mean.anx2, mean.anx.h2, mean.anx.p2)
mean.bel.all2 <- cbind(mean.bel2, mean.bel.h2, mean.bel.p2)

# plot everything
matplot(mean.anx.all2, type="l", ylab="mean anxiety", xlab="days", main="Model 1", col=c("black", "blue", "red"))
legend("bottom", legend = c("total", "healthy", "phobic"), col=c("black", "blue", "red"), pch=16, horiz=T)

matplot(mean.bel.all, type="l", ylab="mean belief", xlab="days", main="Model 1", col=c("black", "blue", "red"))
legend("bottom", legend = c("total", "healthy", "phobic"),  col=c("black", "blue", "red"), pch=16, horiz=T)

matplot(mean.anx.all2, type="l", ylab="mean anxiety", xlab="days", main="Model 2", col=c("black", "blue", "red"))
legend("bottom", legend = c("total", "healthy", "phobic"), col=c("black", "blue", "red"), pch=16, horiz=T)

matplot(mean.bel.all2, type="l", ylab="mean belief", xlab="days", main="Model 2", col=c("black", "blue", "red"))
legend("bottom", legend = c("total", "healthy", "phobic"),  col=c("black", "blue", "red"), pch=16, horiz=T)

```

### Plot VAR model

```{r}

library(vars)
library(qgraph)

# create sequences for both models
sequ.p1 <- seq(from=1, to=nrow(ppl.long.p), by=tpoints)
sequ.h1 <- seq(from=1, to=nrow(ppl.long.h), by=tpoints)
sequ.p2 <- seq(from=1, to=nrow(ppl.long.p2), by=tpoints2)
sequ.h2 <- seq(from=1, to=nrow(ppl.long.h2), by=tpoints2)

# create placeholders 
anx.person.m1 <- matrix(NA, nrow=tpoints, ncol=6)
bel.person.m1 <- matrix(NA, nrow=tpoints, ncol=6)
anx.person.m2 <- matrix(NA, nrow=tpoints2, ncol=6)
bel.person.m2 <- matrix(NA, nrow=tpoints2, ncol=6)

# select 1 healthy vs 1 phobic person
# model 1
vars.p.m1 <- ppl.long[ppl.long[,"ID"]==163,c("anx", "bel")]
v.p.m1 <- VARselect(vars.p.m1, lag.max = 9, type = "const")
vars.h.m1 <- ppl.long[ppl.long[,"ID"]==115,c("anx", "bel")]
v.h.m1 <- VARselect(vars.h.m1, lag.max = 9, type = "const")
# model 2
vars.p.m2 <- ppl.long2[ppl.long[,"ID"]==419,c("anx2", "bel2")]
v.p.m2 <- VARselect(vars.p.m2, lag.max = 9, type = "const")
vars.h.m2 <- ppl.long2[ppl.long[,"ID"]==508,c("anx2", "bel2")]
v.h.m2 <- VARselect(vars.h.m2, lag.max = 9, type = "const")

# lag selection
# model 1
plot(v.p.m1$criteria[1,], type="l", xlab="number of lags", ylab="AIC", main="phobic person model 1")
plot(v.h.m1$criteria[1,], type="l", xlab="number of lags", ylab="AIC", main="healthy person model 1")
# model 2
plot(v.p.m2$criteria[1,], type="l", xlab="number of lags", ylab="AIC", main="phobic person model 2")
plot(v.h.m2$criteria[1,], type="l", xlab="number of lags", ylab="AIC", main="healthy person model 2")

# fit VAR model
# model 1
m.var.p.m1 <- VAR(vars.p.m1, p=1, type = "const")
m.var.h.m1 <- VAR(vars.h.m1, p=1, type = "const")
# model 2
m.var.p.m2 <- VAR(vars.p.m2, p=1, type = "const")
m.var.h.m2 <- VAR(vars.h.m2, p=1, type = "const")

# extract parameters of the VAR model
# model 1
phi.p.m1 <- rbind(coef(m.var.p.m1)$anx[1:2,1], coef(m.var.p.m1)$bel[1:2,1])
phi.h.m1 <- rbind(coef(m.var.h.m1)$anx[1:2,1], coef(m.var.h.m1)$bel[1:2,1])
# model 2
phi.p.m2 <- rbind(coef(m.var.p.m2)$anx[1:2,1], coef(m.var.p.m2)$bel[1:2,1])
phi.h.m2 <- rbind(coef(m.var.h.m2)$anx[1:2,1], coef(m.var.h.m2)$bel[1:2,1])

# make network plots
par(mfrow=c(2,2))
qgraph(t(phi.p.m1), edge.labels = T, layout ="circle", fade = FALSE, labels=c("anxiety", "belief"), negDashed=T, edge.label.cex=1.5)
legend("top", legend = c("Phobic Person Model 1"), horiz=T, bty="n")
qgraph(t(phi.h.m1), edge.labels = T, layout ="circle", fade = FALSE, labels=c("anxiety", "belief"), negDashed=T, edge.label.cex=1.5)
legend("top", legend = c("Healthy Person Model 1"), horiz=T, bty="n")
qgraph(t(phi.p.m2), edge.labels = T, layout ="circle", fade = FALSE, labels=c("anxiety", "belief"), negDashed=T, edge.label.cex=1.5)
legend("top", legend = c("Phobic Person Model 2"), horiz=T, bty="n")
qgraph(t(phi.h.m2), edge.labels = T, layout ="circle", fade = FALSE, labels=c("anxiety", "belief"), negDashed=T, edge.label.cex=1.5)
legend("top", legend = c("Healthy Person Model 2"), horiz=T, bty="n")
```

### Emulate SMSP-A questionnaire

This chunk is for emulating the SMSP-A with model 1 - necessary for later plots
```{r SMSP-A Model 1}

# placeholders
occurrences <- matrix(NA, nrow = nrow(ppl.long), ncol = 11)
colnames(occurrences) <- c("ID", "occ anx 1", "occ anx 2", "occ anx 3", "occ bel 1", "occ bel 2", "occ bel 3", "occ rsb", "occ psb", "occ anxbel", "occ anxrsb")
anx1.occ <- rep(NA, tpoints)
anx2.occ <- rep(NA, tpoints)
anx3.occ <- rep(NA, tpoints)
bel1.occ <- rep(NA, tpoints)
bel2.occ <- rep(NA, tpoints)
bel3.occ <- rep(NA, tpoints)
rsb.occ <- rep(NA, tpoints)
psb.occ <- rep(NA, tpoints)
anxbel.occ <- rep(NA, tpoints)
anxrsb.occ <- rep(NA, tpoints)

# make data for occurrences of symptoms in trigger situations
for(j in sequ){
  for (i in j:(j+tpoints-1)){
    anx1.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,3]>=1))
    anx2.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,3]>=2))
    anx3.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,3]==3))
    bel1.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,5]>=1))
    bel2.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,5]>=2))
    bel3.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,5]==3))
    rsb.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,4]==1))
    psb.occ[i] <- length(which(ppl.long[i,2]==1 && ppl.long[i,6]==1))
    anxbel.occ[i] <- length(which( (ppl.long[i,2]==1 && ppl.long[i,3]>=1) ||                                                    (ppl.long[i,2]==1 && ppl.long[i,5]>=1)))
    anxrsb.occ[i] <- length(which( (ppl.long[i,2]==1 && ppl.long[i,3]>=2) ||                                                    (ppl.long[i,2]==1 && ppl.long[i,4]>=1)))
  }
  occurrences[j:(j+tpoints-1), 1] <- which(sequ==j)
  occurrences[j:(j+tpoints-1), 2] <- anx1.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 3] <- anx2.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 4] <- anx3.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 5] <- bel1.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 6] <- bel2.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 7] <- bel3.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 8] <- rsb.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 9] <- psb.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 10] <- anxbel.occ[j:(j+tpoints-1)]
  occurrences[j:(j+tpoints-1), 11] <- anxrsb.occ[j:(j+tpoints-1)]
}

# create sequence for for loop
sequj <- seq(from=30, to=tpoints, by=30)

# placeholders
total <- rep(NA, length(sequj))
healthy <- rep(NA, length(sequj))
phobic <- rep(NA, length(sequj))

# until what time point should be measured:
tpoints.smsp <- 360

# only take the last seven days 
for(i in 1:nrow(ppl.wide)){
  anx1.occ <- occurrences[occurrences[,1]==i,2]
  ppl.wide$anx1.occ7[i] <- sum(anx1.occ[(tpoints.smsp-6):tpoints.smsp])
  anx2.occ <- occurrences[occurrences[,1]==i,3]
  ppl.wide$anx2.occ7[i] <- sum(anx2.occ[(tpoints.smsp-6):tpoints.smsp])
  anx3.occ <- occurrences[occurrences[,1]==i,4]
  ppl.wide$anx3.occ7[i] <- sum(anx3.occ[(tpoints.smsp-6):tpoints.smsp])
  bel1.occ <- occurrences[occurrences[,1]==i,5]
  ppl.wide$bel1.occ7[i] <- sum(bel1.occ[(tpoints.smsp-6):tpoints.smsp])
  bel2.occ <- occurrences[occurrences[,1]==i,6]
  ppl.wide$bel2.occ7[i] <- sum(bel2.occ[(tpoints.smsp-6):tpoints.smsp])
  bel3.occ <- occurrences[occurrences[,1]==i,7]
  ppl.wide$bel3.occ7[i] <- sum(bel3.occ[(tpoints.smsp-6):tpoints.smsp])
  rsb.occ <- occurrences[occurrences[,1]==i,8]
  ppl.wide$rsb.occ7[i] <- sum(rsb.occ[(tpoints.smsp-6):tpoints.smsp])
  psb.occ <- occurrences[occurrences[,1]==i,9]
  ppl.wide$psb.occ7[i] <- sum(psb.occ[(tpoints.smsp-6):tpoints.smsp])
  anxbel.occ <- occurrences[occurrences[,1]==i,10]
  ppl.wide$anxbel.occ7[i] <- sum(anxbel.occ[(tpoints.smsp-6):tpoints.smsp])
  anxrsb.occ <- occurrences[occurrences[,1]==i,11]
  ppl.wide$anxrsb.occ7[i] <- sum(anxrsb.occ[(tpoints.smsp-6):tpoints.smsp])
}

# placeholders
smsp <- matrix(NA, nrow=nrow(ppl.wide), ncol=7)
colnames(smsp) <- c("i1", "i2", "i3", "i4", "i6", "i7", "i10")

for (i in 1:nrow(ppl.wide)) {
  # item 1
  smsp[i,1] <- if (ppl.wide$anx1.occ7[i] == 0) {0}
               else if (ppl.wide$anx1.occ7[i] <= 2) {1}
               else if (ppl.wide$anx1.occ7[i] <= 4) {2}
               else if (ppl.wide$anx1.occ7[i] <= 6) {3}
               else {4}
  # item 2
  smsp[i,2] <- if (ppl.wide$anxbel.occ7[i] == 0) {0}
               else if (ppl.wide$anxbel.occ7[i] <= 2) {1}
               else if (ppl.wide$anxbel.occ7[i] <= 4) {2}
               else if (ppl.wide$anxbel.occ7[i] <= 6) {3}
               else {4}
  # item 3
  smsp[i,3] <- if (ppl.wide$bel1.occ7[i] == 0) {0}
               else if (ppl.wide$bel1.occ7[i] <= 2) {1}
               else if (ppl.wide$bel1.occ7[i] <= 4) {2}
               else if (ppl.wide$bel1.occ7[i] <= 6) {3}
               else {4}
  # item 4
  smsp[i,4] <- if (ppl.wide$anx2.occ7[i] == 0) {0}
               else if (ppl.wide$anx2.occ7[i] <= 2) {1}
               else if (ppl.wide$anx2.occ7[i] <= 4) {2}
               else if (ppl.wide$anx2.occ7[i] <= 6) {3}
               else {4}
  # item 6
  smsp[i,5] <- if (ppl.wide$psb.occ7[i] == 0) {0}
               else if (ppl.wide$psb.occ7[i] <= 2) {1}
               else if (ppl.wide$psb.occ7[i] <= 4) {2}
               else if (ppl.wide$psb.occ7[i] <= 6) {3}
               else {4}
  # item 7
  smsp[i,6] <- if (ppl.wide$rsb.occ7[i] == 0) {0}
               else if (ppl.wide$rsb.occ7[i] <= 2) {1}
               else if (ppl.wide$rsb.occ7[i] <= 4) {2}
               else if (ppl.wide$rsb.occ7[i] <= 6) {3}
               else {4}
  # item 10
  smsp[i,7] <- if (ppl.wide$anxrsb.occ7[i] == 0) {0}
               else if (ppl.wide$anxrsb.occ7[i] <= 2) {1}
               else if (ppl.wide$anxrsb.occ7[i] <= 4) {2}
               else if (ppl.wide$anxrsb.occ7[i] <= 6) {3}
               else {4}
  ppl.wide$total[i] <- rowSums(smsp)[i] 
}
```

This chunk is for emulating the SMSP-A with model 2 - necessary for later plots

```{r SMSP-A Model 2}
# placeholders
occurrences2 <- matrix(NA, nrow = nrow(ppl.long2), ncol = 11)
colnames(occurrences2) <- c("ID", "occ anx 1", "occ anx 2", "occ anx 3", "occ bel 1", "occ bel 2", "occ bel 3", "occ rsb", "occ psb", "occ anxbel", "occ anxrsb")
anx1.occ2 <- rep(NA, tpoints2)
anx2.occ2 <- rep(NA, tpoints2)
anx3.occ2 <- rep(NA, tpoints2)
bel1.occ2 <- rep(NA, tpoints2)
bel2.occ2 <- rep(NA, tpoints2)
bel3.occ2 <- rep(NA, tpoints2)
rsb.occ2 <- rep(NA, tpoints2)
psb.occ2 <- rep(NA, tpoints2)
anxbel.occ2 <- rep(NA, tpoints2)
anxrsb.occ2 <- rep(NA, tpoints2)

# make data for occurrences of symptoms in trigger situations
for(j in sequ2){
  for (i in j:(j+tpoints2-1)){
    anx1.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,3]>=1))
    anx2.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,3]>=2))
    anx3.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,3]==3))
    bel1.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,5]>=1))
    bel2.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,5]>=2))
    bel3.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,5]==3))
    rsb.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,4]==1))
    psb.occ2[i] <- length(which(ppl.long2[i,2]==1 && ppl.long2[i,6]==1))
    anxbel.occ2[i] <- length(which( (ppl.long2[i,2]==1 && ppl.long2[i,3]>=1) ||                                                  (ppl.long2[i,2]==1 && ppl.long2[i,5]>=1)))
    anxrsb.occ2[i] <- length(which( (ppl.long2[i,2]==1 && ppl.long2[i,3]>=2) ||                                                  (ppl.long2[i,2]==1 && ppl.long2[i,4]>=1)))
  }
  occurrences2[j:(j+tpoints2-1), 1] <- which(sequ2==j)
  occurrences2[j:(j+tpoints2-1), 2] <- anx1.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 3] <- anx2.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 4] <- anx3.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 5] <- bel1.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 6] <- bel2.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 7] <- bel3.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 8] <- rsb.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 9] <- psb.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 10] <- anxbel.occ2[j:(j+tpoints2-1)]
  occurrences2[j:(j+tpoints2-1), 11] <- anxrsb.occ2[j:(j+tpoints2-1)]
}

# create sequence for for loop
sequj2 <- seq(from=30, to=tpoints2, by=30)

# placeholders
total2 <- rep(NA, length(sequj2))
healthy2 <- rep(NA, length(sequj2))
phobic2 <- rep(NA, length(sequj2))

# until what time point should be measured:
tpoints.smsp2 <- 360

# only take the last seven days
for(i in 1:nrow(ppl.wide2)){
  anx1.occ2 <- occurrences2[occurrences2[,1]==i,2]
  ppl.wide2$anx1.occ7[i] <- sum(anx1.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  anx2.occ2 <- occurrences2[occurrences2[,1]==i,3]
  ppl.wide2$anx2.occ7[i] <- sum(anx2.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  anx3.occ2 <- occurrences2[occurrences2[,1]==i,4]
  ppl.wide2$anx3.occ7[i] <- sum(anx3.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  bel1.occ2 <- occurrences2[occurrences2[,1]==i,5]
  ppl.wide2$bel1.occ7[i] <- sum(bel1.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  bel2.occ2 <- occurrences2[occurrences2[,1]==i,6]
  ppl.wide2$bel2.occ7[i] <- sum(bel2.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  bel3.occ2 <- occurrences2[occurrences2[,1]==i,7]
  ppl.wide2$bel3.occ7[i] <- sum(bel3.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  rsb.occ2 <- occurrences2[occurrences2[,1]==i,8]
  ppl.wide2$rsb.occ7[i] <- sum(rsb.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  psb.occ2 <- occurrences2[occurrences2[,1]==i,9]
  ppl.wide2$psb.occ7[i] <- sum(psb.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  anxbel.occ2 <- occurrences2[occurrences2[,1]==i,10]
  ppl.wide2$anxbel.occ7[i] <- sum(anxbel.occ2[(tpoints.smsp2-6):tpoints.smsp2])
  anxrsb.occ2 <- occurrences2[occurrences2[,1]==i,11]
  ppl.wide2$anxrsb.occ7[i] <- sum(anxrsb.occ2[(tpoints.smsp2-6):tpoints.smsp2])
}

# placeholders
smsp2 <- matrix(NA, nrow=nrow(ppl.wide2), ncol=7)
colnames(smsp2) <- c("i1", "i2", "i3", "i4", "i6", "i7", "i10")

for (i in 1:nrow(ppl.wide2)) {
  # item 1
  smsp2[i,1] <- if (ppl.wide2$anx1.occ7[i] == 0) {0}
               else if (ppl.wide2$anx1.occ7[i] <= 2) {1}
               else if (ppl.wide2$anx1.occ7[i] <= 4) {2}
               else if (ppl.wide2$anx1.occ7[i] <= 6) {3}
               else {4}
  # item 2
  smsp2[i,2] <- if (ppl.wide2$anxbel.occ7[i] == 0) {0}
               else if (ppl.wide2$anxbel.occ7[i] <= 2) {1}
               else if (ppl.wide2$anxbel.occ7[i] <= 4) {2}
               else if (ppl.wide2$anxbel.occ7[i] <= 6) {3}
               else {4}
  # item 3
  smsp2[i,3] <- if (ppl.wide2$bel1.occ7[i] == 0) {0}
               else if (ppl.wide2$bel1.occ7[i] <= 2) {1}
               else if (ppl.wide2$bel1.occ7[i] <= 4) {2}
               else if (ppl.wide2$bel1.occ7[i] <= 6) {3}
               else {4}
  # item 4
  smsp2[i,4] <- if (ppl.wide2$anx2.occ7[i] == 0) {0}
               else if (ppl.wide2$anx2.occ7[i] <= 2) {1}
               else if (ppl.wide2$anx2.occ7[i] <= 4) {2}
               else if (ppl.wide2$anx2.occ7[i] <= 6) {3}
               else {4}
  # item 6
  smsp2[i,5] <- if (ppl.wide2$psb.occ7[i] == 0) {0}
               else if (ppl.wide2$psb.occ7[i] <= 2) {1}
               else if (ppl.wide2$psb.occ7[i] <= 4) {2}
               else if (ppl.wide2$psb.occ7[i] <= 6) {3}
               else {4}
  # item 7
  smsp2[i,6] <- if (ppl.wide2$rsb.occ7[i] == 0) {0}
               else if (ppl.wide2$rsb.occ7[i] <= 2) {1}
               else if (ppl.wide2$rsb.occ7[i] <= 4) {2}
               else if (ppl.wide2$rsb.occ7[i] <= 6) {3}
               else {4}
  # item 10
  smsp2[i,7] <- if (ppl.wide2$anxrsb.occ7[i] == 0) {0}
               else if (ppl.wide2$anxrsb.occ7[i] <= 2) {1}
               else if (ppl.wide2$anxrsb.occ7[i] <= 4) {2}
               else if (ppl.wide2$anxrsb.occ7[i] <= 6) {3}
               else {4}
  ppl.wide2$total[i] <- rowSums(smsp2)[i]
}
```


### Correlation Matrices

Now, we can make and visualize correlation matrices of the emulated items for both models
```{r}
corr.smsp <- cor(smsp)
corr.smsp2 <- cor(smsp2)

par(mfrow=c(1,2))
image(corr.smsp, xaxt="n", yaxt = "n", main="Model 1")
image(corr.smsp2, xaxt = "n", yaxt = "n", main="Model 2")
```


### Principal Component Analysis (PCA)

Run a PCA over the items for both models

```{r}
library(ltm)
library(factoextra)
library(multilevel)

# Cronbach's alpha
cronbach.alpha(smsp, CI=T) 
cronbach.alpha(smsp2, CI=T) 

# PCA
pca <- prcomp(smsp, scale=T)
print(pca2)

# item-total correlations
item.total(smsp)
item.total(smsp2)
```

### Load CPES data

The CPES data can be retrieved under https://www.icpsr.umich.edu/web/ICPSR/studies/20240# 
Choose SPSS format for download.

```{r Load CPES Data}
library(foreign)
# set the working directory to downloaded data and name the file "cpes.sav"
dat <- read.spss("cpes.sav")

# Data cleaning: binarise items
i1_CPES <- dat$SP16
i1_CPES <- as.character(i1_CPES)
i1_CPES[i1_CPES != "NOT AT ALL"] <- 1      # Replace values
i1_CPES[i1_CPES == "NOT AT ALL"] <- 0    
i1_CPES <- as.numeric(i1_CPES)

i2_CPES <- dat$SP18 
i2_CPES <- as.character(i2_CPES)
i2_CPES[i2_CPES == "YES"] <- 1             # Replace values
i2_CPES[i2_CPES == "NO"] <- 0
i2_CPES <- as.numeric(i2_CPES)

i3_CPES <- dat$SP22 
i3_CPES <- as.character(i3_CPES)
i3_CPES[i3_CPES != "NEVER"] <- 1           # binarise
i3_CPES[i3_CPES == "NEVER"] <- 0           # Replace values
i3_CPES <- as.numeric(i3_CPES)  

# bind them together
isingCPES_NA <- cbind(i1_CPES, i2_CPES, i3_CPES)
# only use complete cases
isingCPES <- isingCPES_NA[complete.cases(isingCPES_NA),]
```

### Fit Ising Model to the two Models and to CPES Data

```{r Ising Model}
# First emulate measurement of the 3 CPES items with measurement functions
# avoidance = RSB + PSB
ppl.wide$avsum <- ppl.wide$rsbsum + ppl.wide$psbsum
ppl.wide2$avsum2 <- ppl.wide2$rsbsum2 + ppl.wide2$psbsum2

# measurement functions model 1
for(i in 1:N){
ppl.wide$i1[i] <- ifelse(ppl.wide[i, "fearsum"] >= 180 & ppl.wide[i, "avsum"] >= 100 ,1,0)
ppl.wide$i2[i] <- ifelse(ppl.wide[i,"peakanx"] >= 8, 1, 0)
ppl.wide$i3[i] <- ifelse(ppl.wide[i, "avsum"] >= 8, 1, 0)
}
# measurement functions model 2
for(i in 1:N2){
ppl.wide2$i1[i] <- ifelse(ppl.wide2[i, "fearsum2"] >= 180 & ppl.wide2[i, "avsum2"] >= 100 ,1,0)
ppl.wide2$i2[i] <- ifelse(ppl.wide2[i, "peakanx2"] >= 8, 1, 0)
ppl.wide2$i3[i] <- ifelse(ppl.wide2[i, "avsum2"] >= 8, 1, 0)
}

# fit ising models 
res1 <- IsingFit(ising1, family='binomial', plot=T)
res2 <- IsingFit(ising2, family='binomial', plot=T)
resCPES <- IsingFit(isingCPES, family='binomial', plot=T)

# make contingency tables 
# model 1
t12.1 <- table(ppl.wide[,"i1"],ppl.wide[,"i2"])/nrow(ppl.wide)*100
t13.1 <- table(ppl.wide[,"i1"],ppl.wide[,"i3"])/nrow(ppl.wide)*100
t23.1 <- table(ppl.wide[,"i2"],ppl.wide[,"i3"])/nrow(ppl.wide)*100
# model 2
t12.2 <- table(ppl.wide2[,"i1"],ppl.wide2[,"i2"])/nrow(ppl.wide2)*100
t13.2 <- table(ppl.wide2[,"i1"],ppl.wide2[,"i3"])/nrow(ppl.wide2)*100
t23.2 <- table(ppl.wide2[,"i2"],ppl.wide2[,"i3"])/nrow(ppl.wide2)*100
# CPES data
t12.cpes <- table(isingCPES[,"i1_CPES"], isingCPES[,"i2_CPES"])/nrow(isingCPES)*100
t13.cpes <- table(isingCPES[,"i1_CPES"], isingCPES[,"i3_CPES"])/nrow(isingCPES)*100
t23.cpes <- table(isingCPES[,"i2_CPES"], isingCPES[,"i3_CPES"])/nrow(isingCPES)*100

# chi square tests
chisq.test(t12.1)
chisq.test(t12.2)
chisq.test(t12.cpes)

chisq.test(t13.1)
chisq.test(t13.2)
chisq.test(t13.cpes)

chisq.test(t23.1)
chisq.test(t23.2)
chisq.test(t23.cpes)

library(effectsize)
# get phi coefficient for every item pair and every data source
phi(t12.1)
phi(t12.2)
phi(t12.cpes)

phi(t13.1)
phi(t13.2)
phi(t13.cpes)

phi(t23.1)
phi(t23.2)
phi(t23.cpes)

# plot parameters of ising models 
layout(t(1:3))
qgraph(res1$weiadj,fade = FALSE, edge.labels=T, edge.label.cex=2)
title("Model 1")
qgraph(res2$weiadj,fade = FALSE, edge.labels=T, edge.label.cex=2)
title("Model 2")
qgraph(resCPES$weiadj,fade = FALSE, edge.labels=T, edge.label.cex=2)
title("Observed network")
```



